\chapter{Dataset Summary}\label{chap:appendix-1-datasets}

In this appendix we list the variables set discussed in the rest of the document.

\begin{table}[htbp]
    \centering
    \caption[List of 60 time-series channels]{List of 60 time-series channels per session. Each channel corresponds to a specific joint, measurement type (angle or velocity), and axis (X, Y, Z) for left (L) and right (R) sides. \label{tab:met-ts-vars}}
    \begin{tabular}{lll}
    \hline
    \textbf{X-axis Channel} & \textbf{Y-axis Channel} & \textbf{Z-axis Channel} \\
    \hline
    L\_ankle\_angle\_X      & L\_ankle\_angle\_Y      & L\_ankle\_angle\_Z \\
    L\_ankle\_velocity\_X   & L\_ankle\_velocity\_Y   & L\_ankle\_velocity\_Z \\
    L\_foot\_angle\_X       & L\_foot\_angle\_Y       & L\_foot\_angle\_Z \\
    L\_foot\_velocity\_X    & L\_foot\_velocity\_Y    & L\_foot\_velocity\_Z \\
    L\_hip\_angle\_X        & L\_hip\_angle\_Y        & L\_hip\_angle\_Z \\
    L\_hip\_velocity\_X     & L\_hip\_velocity\_Y     & L\_hip\_velocity\_Z \\
    L\_knee\_angle\_X       & L\_knee\_angle\_Y       & L\_knee\_angle\_Z \\
    L\_knee\_velocity\_X    & L\_knee\_velocity\_Y    & L\_knee\_velocity\_Z \\
    R\_ankle\_angle\_X      & R\_ankle\_angle\_Y      & R\_ankle\_angle\_Z \\
    R\_ankle\_velocity\_X   & R\_ankle\_velocity\_Y   & R\_ankle\_velocity\_Z \\
    R\_foot\_angle\_X       & R\_foot\_angle\_Y       & R\_foot\_angle\_Z \\
    R\_foot\_velocity\_X    & R\_foot\_velocity\_Y    & R\_foot\_velocity\_Z \\
    R\_hip\_angle\_X        & R\_hip\_angle\_Y        & R\_hip\_angle\_Z \\
    R\_hip\_velocity\_X     & R\_hip\_velocity\_Y     & R\_hip\_velocity\_Z \\
    R\_knee\_angle\_X       & R\_knee\_angle\_Y       & R\_knee\_angle\_Z \\
    R\_knee\_velocity\_X    & R\_knee\_velocity\_Y    & R\_knee\_velocity\_Z \\
    L\_pelvis\_angle\_X     & L\_pelvis\_angle\_Y     & L\_pelvis\_angle\_Z \\
    L\_pelvis\_velocity\_X  & L\_pelvis\_velocity\_Y  & L\_pelvis\_velocity\_Z \\
    R\_pelvis\_angle\_X     & R\_pelvis\_angle\_Y     & R\_pelvis\_angle\_Z \\
    R\_pelvis\_velocity\_X  & R\_pelvis\_velocity\_Y  & R\_pelvis\_velocity\_Z \\
    \hline
    \end{tabular}
\end{table}


{\small
\setlength{\tabcolsep}{3pt}
\begin{longtable}{@{}p{0.30\textwidth} p{0.60\textwidth} p{0.10\textwidth}@{}}
    \caption{Baseline feature dictionary}\label{tab:baseline_features}\\
    \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Units} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Units} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{3}{r}{\small Continued on next page} \\
    \endfoot
    \bottomrule
    \endlastfoot
    \texttt{step\_width\_left} & Mean step width during left steps (mediolateral distance between foot placements). & m \\
    \texttt{step\_width\_right} & Mean step width during right steps (mediolateral distance between foot placements). & m \\
    \texttt{stride\_rate\_left} & Average stride rate for left steps. & steps/min \\
    \texttt{stride\_rate\_right} & Average stride rate for right steps. & steps/min \\
    \texttt{stride\_length\_left} & Average stride length for left steps (anterior-posterior distance). & m \\
    \texttt{stride\_length\_right} & Average stride length for right steps (anterior-posterior distance). & m \\
    \texttt{swing\_time\_left} & Duration of swing phase for the left limb. & s \\
    \texttt{swing\_time\_right} & Duration of swing phase for the right limb. & s \\
    \texttt{stance\_time\_left} & Duration of stance phase for the left limb. & s \\
    \texttt{stance\_time\_right} & Duration of stance phase for the right limb. & s \\
    \texttt{pelvis\_peak\_drop\_angle\_left} & Peak pelvic obliquity (contralateral drop) during left stance (frontal plane). & $^\circ$ \\
    \texttt{pelvis\_peak\_drop\_angle\_right} & Peak pelvic obliquity (contralateral drop) during right stance (frontal plane). & $^\circ$ \\
    \texttt{pelvis\_drop\_excursion\_left} & Pelvic obliquity range of motion during left stance (frontal plane). & $^\circ$ \\
    \texttt{pelvis\_drop\_excursion\_right} & Pelvic obliquity range of motion during right stance (frontal plane). & $^\circ$ \\
    \texttt{ankle\_df\_peak\_angle\_left} & Peak ankle dorsiflexion angle for the left limb (sagittal plane). & $^\circ$ \\
    \texttt{ankle\_df\_peak\_angle\_right} & Peak ankle dorsiflexion angle for the right limb (sagittal plane). & $^\circ$ \\
    \texttt{ankle\_eve\_peak\_angle\_left} & Peak ankle eversion angle for the left limb (frontal plane). & $^\circ$ \\
    \texttt{ankle\_eve\_peak\_angle\_right} & Peak ankle eversion angle for the right limb (frontal plane). & $^\circ$ \\
    \texttt{ankle\_eve\_percent\_stance\_left} & Timing of peak ankle eversion during left stance, as percent of stance. & \% stance \\
    \texttt{ankle\_eve\_percent\_stance\_right} & Timing of peak ankle eversion during right stance, as percent of stance. & \% stance \\
    \texttt{ankle\_eve\_excursion\_left} & Ankle inversion-eversion range of motion for the left limb (frontal plane). & $^\circ$ \\
    \texttt{ankle\_eve\_excursion\_right} & Ankle inversion-eversion range of motion for the right limb (frontal plane). & $^\circ$ \\
    \texttt{ankle\_rot\_peak\_angle\_left} & Peak ankle axial rotation (internal/external) for the left limb (transverse plane). & $^\circ$ \\
    \texttt{ankle\_rot\_peak\_angle\_right} & Peak ankle axial rotation (internal/external) for the right limb (transverse plane). & $^\circ$ \\
    \texttt{ankle\_rot\_excursion\_left} & Ankle axial rotation range of motion for the left limb (transverse plane). & $^\circ$ \\
    \texttt{ankle\_rot\_excursion\_right} & Ankle axial rotation range of motion for the right limb (transverse plane). & $^\circ$ \\
    \texttt{knee\_flex\_peak\_angle\_left} & Peak knee flexion angle for the left limb (sagittal plane). & $^\circ$ \\
    \texttt{knee\_flex\_peak\_angle\_right} & Peak knee flexion angle for the right limb (sagittal plane). & $^\circ$ \\
    \texttt{knee\_add\_peak\_angle\_left} & Peak knee adduction angle for the left limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_add\_peak\_angle\_right} & Peak knee adduction angle for the right limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_add\_excursion\_left} & Knee adduction-abduction range of motion for the left limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_add\_excursion\_right} & Knee adduction-abduction range of motion for the right limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_abd\_peak\_angle\_left} & Peak knee abduction angle for the left limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_abd\_peak\_angle\_right} & Peak knee abduction angle for the right limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_abd\_excursion\_left} & Knee abduction-adduction range of motion for the left limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_abd\_excursion\_right} & Knee abduction-adduction range of motion for the right limb (frontal plane). & $^\circ$ \\
    \texttt{knee\_rot\_peak\_angle\_left} & Peak knee axial rotation (internal/external) for the left limb (transverse plane). & $^\circ$ \\
    \texttt{knee\_rot\_peak\_angle\_right} & Peak knee axial rotation (internal/external) for the right limb (transverse plane). & $^\circ$ \\
    \texttt{knee\_rot\_excursion\_left} & Knee axial rotation range of motion for the left limb (transverse plane). & $^\circ$ \\
    \texttt{knee\_rot\_excursion\_right} & Knee axial rotation range of motion for the right limb (transverse plane). & $^\circ$ \\
    \texttt{hip\_ext\_peak\_angle\_left} & Peak hip extension angle for the left limb (sagittal plane). & $^\circ$ \\
    \texttt{hip\_ext\_peak\_angle\_right} & Peak hip extension angle for the right limb (sagittal plane). & $^\circ$ \\
    \texttt{hip\_add\_peak\_angle\_left} & Peak hip adduction angle for the left limb (frontal plane). & $^\circ$ \\
    \texttt{hip\_add\_peak\_angle\_right} & Peak hip adduction angle for the right limb (frontal plane). & $^\circ$ \\
    \texttt{hip\_add\_excursion\_left} & Hip adduction-abduction range of motion for the left limb (frontal plane). & $^\circ$ \\
    \texttt{hip\_add\_excursion\_right} & Hip adduction-abduction range of motion for the right limb (frontal plane). & $^\circ$ \\
    \texttt{hip\_rot\_peak\_angle\_left} & Peak hip axial rotation (internal/external) for the left limb (transverse plane). & $^\circ$ \\
    \texttt{hip\_rot\_peak\_angle\_right} & Peak hip axial rotation (internal/external) for the right limb (transverse plane). & $^\circ$ \\
    \texttt{hip\_rot\_excursion\_left} & Hip axial rotation range of motion for the left limb (transverse plane). & $^\circ$ \\
    \texttt{hip\_rot\_excursion\_right} & Hip axial rotation range of motion for the right limb (transverse plane). & $^\circ$ \\
    \texttt{foot\_prog\_angle\_left} & Mean foot progression angle relative to lab (toe-out +, toe-in -), left. & $^\circ$ \\
    \texttt{foot\_prog\_angle\_right} & Mean foot progression angle relative to lab (toe-out +, toe-in -), right. & $^\circ$ \\
    \texttt{foot\_ang\_at\_hs\_left} & Foot pitch angle at heel strike (sagittal), left. & $^\circ$ \\
    \texttt{foot\_ang\_at\_hs\_right} & Foot pitch angle at heel strike (sagittal), right. & $^\circ$ \\
    \texttt{mhw\_exc\_from\_to\_left} & Medial heel whip excursion from toe-off to peak whip, left (transverse). & $^\circ$ \\
    \texttt{mhw\_exc\_from\_to\_right} & Medial heel whip excursion from toe-off to peak whip, right (transverse). & $^\circ$ \\
    \texttt{ankle\_eve\_peak\_vel\_left} & Peak ankle eversion angular velocity, left. & $^\circ$/s \\
    \texttt{ankle\_eve\_peak\_vel\_right} & Peak ankle eversion angular velocity, right. & $^\circ$/s \\
    \texttt{ankle\_rot\_peak\_vel\_left} & Peak ankle axial rotation angular velocity, left. & $^\circ$/s \\
    \texttt{ankle\_rot\_peak\_vel\_right} & Peak ankle axial rotation angular velocity, right. & $^\circ$/s \\
    \texttt{knee\_abd\_peak\_vel\_left} & Peak knee abduction angular velocity, left. & $^\circ$/s \\
    \texttt{knee\_abd\_peak\_vel\_right} & Peak knee abduction angular velocity, right. & $^\circ$/s \\
    \texttt{knee\_add\_peak\_vel\_left} & Peak knee adduction angular velocity, left. & $^\circ$/s \\
    \texttt{knee\_add\_peak\_vel\_right} & Peak knee adduction angular velocity, right. & $^\circ$/s \\
    \texttt{hip\_abd\_peak\_vel\_left} & Peak hip abduction angular velocity, left. & $^\circ$/s \\
    \texttt{hip\_abd\_peak\_vel\_right} & Peak hip abduction angular velocity, right. & $^\circ$/s \\
    \texttt{knee\_rot\_peak\_vel\_left} & Peak knee axial rotation angular velocity, left. & $^\circ$/s \\
    \texttt{knee\_rot\_peak\_vel\_right} & Peak knee axial rotation angular velocity, right. & $^\circ$/s \\
    \texttt{hip\_rot\_peak\_vel\_left} & Peak hip axial rotation angular velocity, left. & $^\circ$/s \\
    \texttt{hip\_rot\_peak\_vel\_right} & Peak hip axial rotation angular velocity, right. & $^\circ$/s \\
    \texttt{pronation\_onset\_left} & Gait cycle percentage when rearfoot pronation begins, left. & \% gait cycle \\
    \texttt{pronation\_onset\_right} & Gait cycle percentage when rearfoot pronation begins, right. & \% gait cycle \\
    \texttt{supination\_timing\_left} & Gait cycle percentage when rearfoot supination begins, left. & \% gait cycle \\
    \texttt{supination\_timing\_right} & Gait cycle percentage when rearfoot supination begins, right. & \% gait cycle \\
    \texttt{hip\_add\_peak\_vel\_left} & Peak hip adduction angular velocity, left. & $^\circ$/s \\
    \texttt{hip\_add\_peak\_vel\_right} & Peak hip adduction angular velocity, right. & $^\circ$/s \\
    \texttt{pelvic\_drop\_peak\_vel\_left} & Peak pelvic obliquity (drop) angular velocity, left stance. & $^\circ$/s \\
    \texttt{pelvic\_drop\_peak\_vel\_right} & Peak pelvic obliquity (drop) angular velocity, right stance. & $^\circ$/s \\
    \texttt{vertical\_oscillation\_left} & Vertical oscillation amplitude during left stance. & mm \\
    \texttt{vertical\_oscillation\_right} & Vertical oscillation amplitude during right stance. & mm \\
    \texttt{age} & Subject age. & years \\
    \texttt{height} & Body height. & cm \\
    \texttt{weight} & Body weight. & kg \\
    \texttt{gender\_female} & Indicator: female (1) or not (0). & 0/1 \\
    \texttt{gender\_male} & Indicator: male (1) or not (0). & 0/1 \\
    \texttt{dominantleg\_left} & Indicator: dominant leg is left (1) or not (0). & 0/1 \\
    \texttt{dominantleg\_right} & Indicator: dominant leg is right (1) or not (0). & 0/1 \\
\end{longtable}
}

{\small
\setlength{\tabcolsep}{3pt}
\begin{longtable}{@{}p{0.30\textwidth} p{0.60\textwidth} p{0.10\textwidth}@{}}
    \caption{Dominant leg derived feature dictionary}\label{tab:domleg_features}\\
    \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Units} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Units} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{3}{r}{\small Continued on next page} \\
    \endfoot
    \bottomrule
    \endlastfoot
    \texttt{age} & Subject age. & years \\
    \texttt{height} & Body height. & cm \\
    \texttt{weight} & Body weight. & kg \\
    \texttt{dom\_leg\_ankle\_df\_peak\_angle} & Peak ankle dorsiflexion angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_df\_peak\_angle} & Difference in peak ankle dorsiflexion (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_eve\_excursion} & Ankle inversion-eversion range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_eve\_excursion} & Difference in ankle inversion-eversion ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_eve\_peak\_angle} & Peak ankle eversion angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_eve\_peak\_angle} & Difference in peak ankle eversion (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_eve\_peak\_vel} & Peak ankle eversion angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_ankle\_eve\_peak\_vel} & Difference in peak ankle eversion velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_ankle\_eve\_percent\_stance} & Timing of peak ankle eversion in stance, dominant limb. & \% stance \\
    \texttt{dom\_leg\_diff\_ankle\_eve\_percent\_stance} & Difference in timing of peak eversion in stance (dominant minus non-dominant). & \% stance \\
    \texttt{dom\_leg\_ankle\_rot\_excursion} & Ankle axial rotation range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_rot\_excursion} & Difference in ankle axial rotation ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_rot\_peak\_angle} & Peak ankle axial rotation (internal/external), dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_rot\_peak\_angle} & Difference in peak ankle axial rotation (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_rot\_peak\_vel} & Peak ankle axial rotation angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_ankle\_rot\_peak\_vel} & Difference in peak ankle axial rotation velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_foot\_ang\_at\_hs} & Foot pitch angle at heel strike, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_foot\_ang\_at\_hs} & Difference in foot pitch at heel strike (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_foot\_prog\_angle} & Mean foot progression angle relative to laboratory, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_foot\_prog\_angle} & Difference in foot progression angle (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_abd\_peak\_vel} & Peak hip abduction angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_hip\_abd\_peak\_vel} & Difference in peak hip abduction velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_hip\_add\_excursion} & Hip adduction-abduction range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_add\_excursion} & Difference in hip adduction-abduction ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_add\_peak\_angle} & Peak hip adduction angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_add\_peak\_angle} & Difference in peak hip adduction (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_add\_peak\_vel} & Peak hip adduction angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_hip\_add\_peak\_vel} & Difference in peak hip adduction velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_hip\_ext\_peak\_angle} & Peak hip extension angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_ext\_peak\_angle} & Difference in peak hip extension (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_rot\_excursion} & Hip axial rotation range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_rot\_excursion} & Difference in hip axial rotation ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_rot\_peak\_angle} & Peak hip axial rotation, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_rot\_peak\_angle} & Difference in peak hip axial rotation (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_rot\_peak\_vel} & Peak hip axial rotation angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_hip\_rot\_peak\_vel} & Difference in peak hip axial rotation velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_knee\_abd\_excursion} & Knee abduction-adduction range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_abd\_excursion} & Difference in knee abduction-adduction ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_abd\_peak\_angle} & Peak knee abduction angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_abd\_peak\_angle} & Difference in peak knee abduction (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_abd\_peak\_vel} & Peak knee abduction angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_knee\_abd\_peak\_vel} & Difference in peak knee abduction velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_knee\_add\_excursion} & Knee adduction-abduction range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_add\_excursion} & Difference in knee adduction-abduction ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_add\_peak\_angle} & Peak knee adduction angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_add\_peak\_angle} & Difference in peak knee adduction (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_add\_peak\_vel} & Peak knee adduction angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_knee\_add\_peak\_vel} & Difference in peak knee adduction velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_knee\_flex\_peak\_angle} & Peak knee flexion angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_flex\_peak\_angle} & Difference in peak knee flexion (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_rot\_excursion} & Knee axial rotation range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_rot\_excursion} & Difference in knee axial rotation ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_rot\_peak\_angle} & Peak knee axial rotation, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_rot\_peak\_angle} & Difference in peak knee axial rotation (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_rot\_peak\_vel} & Peak knee axial rotation angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_knee\_rot\_peak\_vel} & Difference in peak knee axial rotation velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_mhw\_exc\_from\_to} & Medial heel whip excursion from toe-off to peak whip, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_mhw\_exc\_from\_to} & Difference in medial heel whip excursion (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_pelvic\_drop\_peak\_vel} & Peak pelvic obliquity (drop) angular velocity, dominant stance. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_pelvic\_drop\_peak\_vel} & Difference in pelvic drop peak velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_pelvis\_drop\_excursion} & Pelvic obliquity range of motion during dominant stance. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_pelvis\_drop\_excursion} & Difference in pelvic obliquity ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_pelvis\_peak\_drop\_angle} & Peak pelvic drop angle during dominant stance. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_pelvis\_peak\_drop\_angle} & Difference in pelvic peak drop angle (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_pronation\_onset} & Gait cycle percentage when rearfoot pronation begins, dominant limb. & \% gait cycle \\
    \texttt{dom\_leg\_diff\_pronation\_onset} & Difference in pronation onset timing (dominant minus non-dominant). & \% gait cycle \\
    \texttt{dom\_leg\_stance\_time} & Duration of stance phase, dominant limb. & s \\
    \texttt{dom\_leg\_diff\_stance\_time} & Difference in stance time (dominant minus non-dominant). & s \\
    \texttt{dom\_leg\_step\_width} & Mean step width measured for steps of dominant limb. & m \\
    \texttt{dom\_leg\_stride\_length} & Mean stride length for dominant limb strides. & m \\
    \texttt{dom\_leg\_diff\_stride\_length} & Difference in stride length (dominant minus non-dominant). & m \\
    \texttt{dom\_leg\_stride\_rate} & Stride rate for dominant limb strides. & steps/min \\
    \texttt{dom\_leg\_diff\_stride\_rate} & Difference in stride rate (dominant minus non-dominant). & steps/min \\
    \texttt{dom\_leg\_supination\_timing} & Gait cycle percentage when rearfoot supination begins, dominant limb. & \% gait cycle \\
    \texttt{dom\_leg\_diff\_supination\_timing} & Difference in supination onset timing (dominant minus non-dominant). & \% gait cycle \\
    \texttt{dom\_leg\_swing\_time} & Duration of swing phase, dominant limb. & s \\
    \texttt{dom\_leg\_diff\_swing\_time} & Difference in swing time (dominant minus non-dominant). & s \\
    \texttt{dom\_leg\_vertical\_oscillation} & Vertical oscillation amplitude during dominant stance. & mm \\
    \texttt{dom\_leg\_diff\_vertical\_oscillation} & Difference in vertical oscillation (dominant minus non-dominant). & mm \\
    \texttt{gender\_female} & Indicator: female (1) or not (0). & 0/1 \\
    \texttt{gender\_male} & Indicator: male (1) or not (0). & 0/1 \\
\end{longtable}
}


{\small
\setlength{\tabcolsep}{3pt}
\begin{longtable}{@{}p{0.30\textwidth} p{0.60\textwidth} p{0.10\textwidth}@{}}
    \caption{Top 40 selected feature dictionary}\label{tab:top40_features}\\
    \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Units} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Variable} & \textbf{Description} & \textbf{Units} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{3}{r}{\small Continued on next page} \\
    \endfoot
    \bottomrule
    \endlastfoot
    \texttt{age} & Subject age. & years \\
    \texttt{weight} & Body weight. & kg \\
    \texttt{dom\_leg\_ankle\_df\_peak\_angle} & Peak ankle dorsiflexion angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_eve\_peak\_angle} & Peak ankle eversion angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_eve\_peak\_angle} & Difference in peak ankle eversion (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_eve\_peak\_vel} & Peak ankle eversion angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_ankle\_eve\_percent\_stance} & Difference in timing of peak eversion in stance (dominant minus non-dominant). & \% stance \\
    \texttt{dom\_leg\_ankle\_rot\_excursion} & Ankle axial rotation range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_ankle\_rot\_excursion} & Difference in ankle axial rotation ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_ankle\_rot\_peak\_vel} & Peak ankle axial rotation angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_foot\_ang\_at\_hs} & Foot pitch angle at heel strike, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_foot\_ang\_at\_hs} & Difference in foot pitch at heel strike (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_add\_excursion} & Hip adduction-abduction range of motion, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_hip\_add\_peak\_angle} & Peak hip adduction angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_add\_peak\_angle} & Difference in peak hip adduction (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_add\_peak\_vel} & Peak hip adduction angular velocity, dominant limb. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_hip\_add\_peak\_vel} & Difference in peak hip adduction velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_hip\_ext\_peak\_angle} & Peak hip extension angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_ext\_peak\_angle} & Difference in peak hip extension (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_hip\_rot\_excursion} & Hip axial rotation ROM, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_hip\_rot\_peak\_vel} & Difference in peak hip axial rotation velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_knee\_add\_excursion} & Difference in knee adduction-abduction ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_add\_peak\_vel} & Difference in peak knee adduction velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_knee\_flex\_peak\_angle} & Peak knee flexion angle, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_rot\_excursion} & Difference in knee axial rotation ROM (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_knee\_rot\_peak\_angle} & Peak knee axial rotation, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_knee\_rot\_peak\_vel} & Difference in peak knee axial rotation velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_mhw\_exc\_from\_to} & Medial heel whip excursion from toe-off to peak whip, dominant limb. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_mhw\_exc\_from\_to} & Difference in medial heel whip excursion (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_pelvic\_drop\_peak\_vel} & Peak pelvic obliquity (drop) angular velocity, dominant stance. & $^\circ$/s \\
    \texttt{dom\_leg\_diff\_pelvic\_drop\_peak\_vel} & Difference in pelvic drop peak velocity (dominant minus non-dominant). & $^\circ$/s \\
    \texttt{dom\_leg\_pelvis\_peak\_drop\_angle} & Peak pelvic drop angle during dominant stance. & $^\circ$ \\
    \texttt{dom\_leg\_diff\_pelvis\_peak\_drop\_angle} & Difference in pelvic peak drop angle (dominant minus non-dominant). & $^\circ$ \\
    \texttt{dom\_leg\_stance\_time} & Duration of stance phase, dominant limb. & s \\
    \texttt{dom\_leg\_step\_width} & Mean step width measured for steps of dominant limb. & m \\
    \texttt{dom\_leg\_stride\_length} & Mean stride length for dominant limb strides. & m \\
    \texttt{dom\_leg\_stride\_rate} & Stride rate for dominant limb strides. & steps/min \\
    \texttt{dom\_leg\_supination\_timing} & Gait cycle percentage when rearfoot supination begins, dominant limb. & \% gait cycle \\
    \texttt{dom\_leg\_swing\_time} & Duration of swing phase, dominant limb. & s \\
    \texttt{dom\_leg\_diff\_vertical\_oscillation} & Difference in vertical oscillation amplitude (dominant minus non-dominant). & mm \\
\end{longtable}
}
