\chapter{Conclusions}\label{chap:conclusions}

\section{Conclusions}\label{sec:conc-conclusions}
-- Linked to abstract, very focused.
... baseline models remain a robust choice for injury status classification on this dataset.

\section{Limitations}\label{sec:conc-limitations}
- Normalised Time-series we have lost inter-limb coordination cues. For the model it looks like both legs move at the same time. If there is critical informaiton hiden in how the body reacts when both sides of the body work together, this is lost. A different appraoch would have been to encode a temporal variable that would go from start of one side to the end of the other, the challenges: What is both sides are not simetric? (btw also lost due to segmentation and normalisation), etc..

\subsection{Data limitations (sample, biases, missingness)}\label{sec:limitations-data}

\subsection{Methodological limitations (assumptions, generalization)}\label{sec:limitations-methods}
- Validation to Test score drop (0.1) in deep learning models suggests that validation guided model selection and the test set is truly unseen. An easy improvement is to use CV-based tuning exclusively with the traning set.

\subsection{Evaluation limitations (metrics, external validation)}\label{sec:limitations-evaluation}
-- Approaching the problem as a prediction of the mayority class -> Class imbalnce made things harder
    -- In the future, change to minority class

\section{Future Work}\label{sec:conc-future-work}
-- All of the work that we did not have time to do.
-- What we would do if we had more time.
-- What we would do if we had more data.
-- What we would do if we had more compute.
-- Look into other parts of the gait cycle? Or use the full cycle isntead of only the stance.
-- curver summarization in 3D, not only 1D to preserve better the shape.
-- Maybe try the propblem in 2D?
-- analysing the saliency maps by injury type can help undersantd difference between injuries. for example: Knee injuries will probably havea difference map that fascitic plantar, because the respective parts of the body are exercises and strained during different phases.
-- Run permutation importance on deeplearning models, but making sure permutatiosn are done channel-wise.
